---
title: "GPDI517: Introduction to dplyr"
author: "by Tugba Ozturk"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this tutorial, we will use **mtcars** data frame to go over the following functions in `dplyr`:

- select()
- rename()
- arrange()
- mutate() and transmutate()
- filter()
- group_by() and summarise()


## 1. Install and load the packge!

Let us start by installing and loading the `dplyr` library. To do that, we need to uncomment the lines of code in the following code chunk and run the code: a) we can click on the **Knit** button to run the code and regenerate the document or b) we can click on the **Run Current Chunk** button on the top-right shown with a green arrow. 

```{r install-load-dplyr}

#install.packages("dplyr") #to install the library from the CRAN
#library("dplyr") #to activate/load the library on your workspace, every time you start a new session in R, you should use this command to load the dplyr library!

```

## 2. Get familiar with the data set -- mtcars.

Using `knitr::kable()`, we can generate a table to look at the first 10 lines of the **mtcars** data set. If we wish to see the whole data, we can replace **head(mtcars,10)** with **mtcars** below and **knit** the document: 

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

It is a good idea to look at the summary of a data frame:

```{r summary, echo=FALSE }
if (sum(is.na(mtcars))==0) {
  cat("\nThere is",sum(is.na(mtcars)), "missing data points in this data frame!\n\n")
} else {
  cat("\nNo missing data points!\n")
}

print(summary(mtcars))

```


## 3. Data manipulation with **dplyr**

Write an R script (named **myanswer.R**) to do the following:

- create a local copy of **mtcars**
- select the following columns: mpg, cyl, disp, hp, drat, wt 
- arrange the data frame by the disp column in the descending order
- create a new column which is equal to mpg*cyl/wt and name it x
- group the data by cyl, calculate the average, standard deviation of disp column
and the number of columns, write the output as *dplyr-exercise-disp.txt*
- add a new column which is the copy of cyl column where 4,6, and 8 are represented as "A","B", and "C", respectively
- rename x column as mcw
- create a new data frame with all the rows having a value between 3 and 100 in the disp column and save the new data frame as *dplyr-exercise.out* (white space separated, without quotation marks)


```{r dplyrexercise }
# let's run the script by umcommenting the following line and running the current
# chunk: 
#source("myanswer.R")

```


## 4. **%>%** operator makes it more readable!

Now, copy the content of your script below and rewrite it using the pipe operator (**%>%**):


```{r dplyrexercisewithpipe }


```



## 5. Bonus!

Add your comments to the following code:

```{r twotable }
#df <- data.frame(cyl=c(4,6,8),legend=c("four", "six","eight"))
#a <- left_join(mtcars,df)
```





